### 2017-03-07 resource/index.Rmd
added to resource

---
title: "Understanding GDP relations to Crime in US Cities"
author: "Kai Yoshino, Isabella Spaletta, Chrisopther Ross, Marisa Nanakul"
date: "March 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
source("resources.R")
```

### What Are The Datasets?

The two datasets that we used in this anaylsis are one for crime based data on U.S. cities over a 5 year time period (2010-2015) and one for the GDP for metropolitan areas in the U.S. over a 5 year time period (2010-2015).

#### FBI Crime Dataset

Below is an example of what the crime dataset looks like (shows all 5 years):

```{r, echo=FALSE}
  kable(cities.all[1:5,])
```

##### Observations Made About the Dataset
* The size of the dataset is {rows, columns}:

```{r, echo=FALSE}
  dim(cities.all)
```

* Column Names and Data Types of the crime table:

```{r, echo=FALSE}
  colnames(cities.all)
```

```{r, echo=FALSE}
  sapply(cities.all, typeof)
```

##### Other Observations & Changes Made to Dataset
* All crime datasets were individually captured by year, but we combined them to make one large crime dataset.
* If a city has a crime value of NA, this means this data could not be found, not that any of these crimes did not occur. NA does not equal 0 in this dataset.
* Violent crimes is defined of four offenses: murder and nonnegligent manslaughter, rape, robbery, and aggravated assault.
* Property crimes is defined of four offenses: burglary, larceny theft, motor vehicle theft, and arson.
<br>
<br>

#### GDP Dataset

```{r, echo=FALSE}
  gdp[1:5,] %>% 
  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015") %>% 
  kable()
```

##### Observations Made About the Dataset
* The size of the dataset is {rows, columns}:

```{r, echo=FALSE}
  dim(gdp %>% 
  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
```

* Column Names and Data Types of the crime table:

```{r, echo=FALSE}
  colnames(gdp %>% 
  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
```

```{r, echo=FALSE}
  sapply(gdp %>% 
  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"), typeof)
```

##### Other Observations & Changes Made to Dataset
* All of this GDP data was added to the entire crime dataset.
* We included the GDP each of the five years and GDP per person for each of the five years.
<br>
<br>