### 2017-03-08 paper/paper.Rmd

---
title: "Final Project Paper"
output:
  html_document: default
  html_notebook: default
---

#### Isabella Spaletta, Marisa Nanakul, Christopher Ross, Kai Yoshino
##### Affiliation: University of Washington - Information School

### Abstract

********** Marisa
The key findings of our research showed that a poisson regression showed no correlation between the GDP per person and crime per person in each metropolitan area with population as another independent variable. Between all of the years from 2010 to 2015, there was no significance between GDP per person, crime per person, and population of a metropolitan area for any compared years. From this, our contribution to society and information in general is that just looking at a metropolitan area's economy and population is not a good determining factor on the number of crimes permitted by person.

### Introduction

Currently, there are many different sites and organizations that collect data on cities or large metropolitan areas. The Federal Bureau of Investigation (FBI) collects crime data and population every year for individual cities within the United States and the Bureau of Economic Analysis (BEA) gathered an overall metropolitan area Gross Domestic Product (GDP) every year since 2010. There are a few different organizations or articles that have looked at a comparison between crime rates and the GDP of specific metropolitan areas, but our specific analysis considers population as another independent variable when comparing crime rate and GDP of metropolitan areas. Some of the related work we will discuss takes the change in economy after the 2008 financial crisis into account. One other research group took into account economic uncertainty whereas we will focus more on the count data of a metropolitan area (population, GDP). We want to look at the GDP and crime relationship affected by population to see if GDP per person and population of a metropolitan area have any sort of impact on the crimes per person committed within that metropolitan area.

Our analysis focuses on violent and property crime impact on city GDPs with population as another independent variable. The analysis is a unique contribution because rather than analyzing from the state level, we are choosing to focus on the city level, adding an extra area of granularity. In terms of covariates, we are accounting for the population change between years along with the changes in GDP and crime. We are observing and comparing violent and property crime trends to metropolitan area GDP and population trends.

### Related Work

In the past, most research has been focused around states in the U.S. and countries around the world. In “Falling Together” [1], data from the U.S. Department of Justice from 2007-2010 was used to analyze the relationship between property crime rates and GDP per person in all 50 states. The results found that property crime rates and GDP both decreased during this time frame. Specifically after the recession, the general trend was as GDP decreased so did property crimes. The cause of the trend is unknown.

“The Puzzling Relationship Between Crime and the Economy” [2] delves into the different conclusions economists and criminologists conclude about the relationship between crime and the economy. Criminologists believe that tough economic times increase the likelihood of a person committing a crime. They also believe that economic anxieties cause people to consume more mind-altering substances, leading to increased violence. Economists believe better economic conditions increase the likelihood of a person committing a crime. With technology becoming more visible and ingrained into people’s lives, criminals are able to see what’s possible to steal. Analysis on violent and property crimes from 1960-2012 suggests that there is no relationship between GDP and crime, and that crime is episodic.

In “Economic growth and crime: does uncertainty matter?” [3] 25 countries (including the U.S.) are analyzed with a macroeconomic uncertainty metric to observe whether there is a relationship between crime and GDP. The researchers found that if GDP uncertainty is negative, the crime uncertainty is positive, meaning that crime uncertainty negatively impacts GDP uncertainty.

### Data Prep

The preparation of our data, specifically combining the crime and GDP datasets, was a very important and large aspect of conducting this research. The data given to us was very different in formatting and was therefore fairly difficult to combine. The FBI crime dataset had city names in one column rather than metropolitan areas (some cities combined), and in another column the state names were fully written out instead of having state abbreviations. All of the crime datasets were separated by year equalling six different datasets, whereas the GDP dataset had columns of GDP data separated by year equalling only one dataset that all had to be combined.

To begin combining these two datasets, we joined all of the crime datasets together so that every year was accounted for in each city within the dataset along with the population of each city. From here, we decided that since the GDP dataset had less cities/metropolitan areas to account for, we would take all of the corresponding cities from the crime dataset and add the GDP for each of those cities to the crime dataset. We then removed any city within the crime dataset that did not include a GDP (was not included in the GDP dataset). New rows were added for the metropolitan areas that included more than one city (e.g. Seattle-Tacoma-Bellevue) and added up the population and crimes in each of the cities to get a total for the new row added. This is important because to keep the GDP accurate, we had to keep the cities combined as a metropolitan area, but we could account for the population and crimes by adding the counts together and doing analysis from the new totals. We also removed any metropolitan area from the GDP dataset that included multiple states and had to change any value in the data set that was NA to 0 to be able to combine these datasets properly.

Our final dataset came out as a final CSV that included any city or metropolitan areas that were included in both the crime and GDP dataset. For each of these cities or metropolitan areas, we included the population of each metropolitan area, crime data for 2010-2015, and the GDP for 2010-2015. This came out to include 366 cities or metropolitan areas. We also added columns such as crime per person in each metropolitan area and gdp per person in each metropolitan area to run our analyses on.

One thing to note is that to reproduce this research you will have to use the CSV we created or any other researcher could find other data sources for GDP and crime for the years 2010 to 2015 and create their own new dataset. You must also change any NA in the datset to 0 to combine the data.

Another thing to note is that 
'violent crimes' is defined by four offenses: murder and nonnegligent manslaughter, rape, robbery, and aggravated assault. 
'property crimes' is defined by four offenses: burglary, larceny theft, motor vehicle theft, and arson.

### Analysis

### Results

### Discussion

### Future Work

While we looked at GDP and crime rate, to extend the work we have done here, other researchers can also compare GDP and crime rate per city with other independent variables other than population that may also affect the correlation. One example is other related work looks at economic uncertainty as an independten variable and there are other covariates that could possibly show a correlation between the GDP and crime rate of a city such as education, population density, health metrics, and many more.

One other aspect to consider is the particular datasets used and how the data is measured. Since we are doing by metropolitan area, crime rate to us is defined as total number of crimes divided by the population of each city or metropolitan area within the dataset. The GDP dataset that we chose to use was in a very unfortunate format when trying to combine it with the crime dataset, so I would suggest that in future work an alternative GDP dataset is used that is by actual city in the US rather than metropolitan area.

### References

[3] Goulas, Eleftherios, and Athina Zervoyianni. "Economic Growth and Crime: Does Uncertainty Matter?" Applied Economics Letters 20.5 (2013): 420-27. Web.

[2] Roman, John. "The Puzzling Relationship Between Crime and the Economy." CityLab. The Atlantic, 23 Sept. 2013. Web. 05 Mar. 2017.

[1] The Economist Online. "Falling Together." The Economist. The Economist Newspaper, 26 Sept. 2011. Web. 05 Mar. 2017.